Controlling the dewetting morphologies of thin liquid films by switchable substrates

### 2022-12-06 Reply_RefereeA.tex

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee A}
We thank the Referee for his/her report on our manuscript. 
We are glad to read that he/she finds it of potential interest and in principle 
suitable for publication, if a thorough revision is performed. 
We substantially improved and revised the paper for the resubmission and hope that we were able to clarify the obscure points, following all the Referee's remarks.
Hereafter, we provide answers to the various questions/comments raised. All changes in the revised version are reported in red color.

@@

\item[ \textbf{\underline{Comment 2.}}] 
{
The same holds for the time scale. 
The viscous time scale is well defined, but necessitates a length scale in addition to the 
capillary velocity. It is unclear why the authors choose the length scale that they use in the paper.
Subsequently, they replace the surface tension by a complicated partial derivative of the 
disjoining pressure; here the same remark as above holds.
}

\item[ \textbf{{Answer}}]
{
We apologize if we have not been thorough in the discussion of the relevant scales, which indeed deserves some clarification. 
Thin film hydrodynamics is an intrinsically multiscale phenomenon, whereby there are at least two characteristic length scales: the mean film thickness (let us call it $h_0$)
and a horizontal length scale $\ell$ (which in our case can be, for example, the pattern wavelength, $\lambda$), where, typically $h_0 \ll \ell$. We chose to use the mean 
film thickness as a scale, because it is a relevant length that is common to all the simulations.
In the same spirit, we chose as a reference time $t_0(\theta)$, which is not exactly the capillary time (though being related to it). 
$t_0$, as we explain in the text, is, in fact, the characteristic time of growth of the most unstable mode in the spinodal dewetting on a homogeneous
substrate of contact angle $\theta$. In spite of the substrate being patterned in our case, it is possible to define an effective $t_0$ in terms 
of, e.g., the mean contact angle on the pattern, i.e. $t_0(\theta_0)$. 
We deem the so defined characteristic time as an appropriate scale,
since it does not depend on the pattern wavelength and speed. Actually, one can see that it captures well the time scales of the dewetting process
from figure 2 of the manuscript, where it is shown that the rupture times for the largest wavelength (which are closer to the homogeneous case) are of the order of $t_0$.
We recognize, on the other hand that in this perspective the choice of $v_0 = \lambda/t_0$ as a reference velocity looks a bit awkward.
A more natural candidate would be, indeed, $v_0=\lambda_s/t_0$, where $\lambda_s$ is the spinodal 
characteristic length corresponding to $\theta_0$. We have therefore decided to use the latter in the revised version of the manuscript.
We have also rephrased the corresponding paragraph in the {\it Method} section as follows: \\

@@

\textcolor{red}{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
$q_0^2=h_{\ast}^{-2}(1-\cos\theta_0)f^{\prime}(h_0/h_{\ast})$.
Correspondingly, we choose as a velocity scale $v_0 = \lambda_s/t_0$, where $\lambda_s = 2\pi/q_0$.}

}

\item[ \textbf{\underline{Comment 3.}}]
{ 
Incidentally, for a 'normal' fluid such as water, the capillary velocity is $\sim 70\mbox{m}/\mbox{s}$, 
so that very, very rapidly the fluid inertia becomes important. 
This is indeed what one usually observes for the formation or coalescence of drops, so that at the very least this should be discussed and the Reynolds number calculated.
}

\item[ \textbf{{Answer}}]
{
Such a high value corresponds to the ratio of (surface tension)/(dynamic viscosity) for water, i.e. $v = \gamma/\eta \approx 70 \mbox{m}/\mbox{s}$.
This expression, though, is only representative of the capillary velocity, again, for {\it free surface flows}, as, e.g., in the coalescence of 
two droplets in the bulk of another fluid. 
In the case wetting/dewetting, 
the phenomenology is complicated by the liquid/substrate (and vapour/substrate) interaction. 
In such a situation the relevant velocity scale is set by the characteristic velocity of contact line 
motion which can be estimated from a viscocapillary balance as 
$\sim \varepsilon^3 \gamma/\eta$, where $\varepsilon = h_0/L \ll 1$ is the film thickness parameter 
\cite{RevModPhys.69.931,doi:10.1146/annurev-fluid-011212-140734}. 
The typical velocities in dewetting problems are, then, much smaller than the ratio $\gamma/\eta$.
Moreover, in many actual applications, as well as in the majority of experiments, the 
working fluids are in general much more viscous than water (e.g., polymeric liquids), thus making
the characteristic velocity even smaller.
For the dewetting of a millimetric film of micrometric thickness reatracting into a single droplet, for instance, experiments have measured that 
the typical velocity is $\sim 1 \mbox{mm}/\mbox{s}$, which gives a Reynolds number $Re \approx 10^{-3}$ \cite{Edwards2016}. 
In our problem, there are two relevant characteristic velocities: the retraction speed, 
$U_{\Theta}$, and 
the pattern speed, $v_{\theta}$. When their ratio $\Gamma = v_{\theta}/U_{\Theta} < 1$, the retraction speed dominates and fixes the Reynolds number which can be, then, evaluated as 
$Re = U_{\Theta}h_0/\nu$ \cite{RevModPhys.69.931}, whereas for $\Gamma > 1$, $Re = v_{\theta} h_0/\nu$. 
In any case, the Reynolds number stays always below $\approx 0.2$.
We have added the following comment at the end of the {\it Method} section:\\

@@

Concerning the values of viscosity and surface tension in lattice Boltzmann units, 
we acknowledge that giving them in the middle of the discussion on characteristic length and time might  
have caused confusion. Nevertheless we deem important to provide them, for the sake of readers interested 
in numerical details of the method as well as for the results reproducibility, therefore we have moved them 
to a later paragraph, where we added: \\
\\
\textcolor{red}{The numerical values in lattice Boltzmann units of the 
mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
and $\gamma=0.01$.}
}

\item[ \textbf{\underline{Comment 6.}}]
{ 
What is the disjoining pressure due to? 
Van der Waals forces? 
How thick is the precursor film, and why should the disjoining pressure be zero there?
Is there any comparison with experiments possible/feasible?
}

\item[ \textbf{{Answer}}]
{
As we have already discussed in the first answer (and as one can find 
in several reviews and undergraduate textbooks, but see, e.g., Craster 
and Matar \cite{craster2009dynamics}), the disjoining pressure encodes 
any kind of interactions among the fluid molecules and the substrate, so 
also those of Van der Waals type. The precursor film is an equilibrium value for the thickness that emerges when the interface potential has also a short 
range repulsive term. The disjoining pressure is zero there by definition of equilibrium (or, equivalently, because that is a minimum of the interface potential). Concerning comparisons of a thin-film equation approach with 
experiments, there are tons of examples in the literature, such as 
\cite{becker2003complex,fetzer2005new,PhysRevLett.99.114503} to name but a few (see also \cite{craster2009dynamics,RevModPhys.81.739} for reviews). All this literature is cited in the paper.
}

### 2022-12-06 Reply_RefereeB.tex

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee B}

We thank the Referee for his/her thorough reading of our manuscript and detailed report.
We are happy to read that the referee acknowledges the amount of work we did and thinks it could be fitting for PRL after revision.
In what follows we provide detailed answers to all Referee's questions and we discuss all the changes made to the manuscript
(which are typeset in red color).

@@

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{
The model disjoining pressure, Eq. 3 with a tail decaying as the
inverse squared power seems odd. I am more familiar with the expected
van der Waals tail with a decay as the inverse cube power. Is there
any reason for this choice?

\item[ \textbf{Answer}]
{
The Referee is right in pointing out that, indeed, the $h^{-3}$ decay 
for the attractive part of the disjoining pressure is more straightforward 
to understand as it can be derived directly from an intermolecular potential of Lennard-Jones type. Nevertheless, it is also true that other forms
can be encountered in the literature, depending on the actual interactions
\cite{SCHWARTZ1998173,Mitlin,Teletzke}. Here, though, our choice of the $(3,2)$ pair of exponents is motivated more by computational reasons rather than by the will of 
simulating a specific liquid with given physical properties.
With such choice, in fact, the characteristic dewetting time 
is shorter, thus making the overall computational cost more manageable.
In fact, we have seen in a previous study~\cite{PhysRevE.104.034801} that,
except for this shift of the time scales, the dewetting dynamics remains 
qualitatively the same. To confirm this observation, we have run 
a new simulation (for a single pair of pattern wavelength and speed) 
with the $(9,3)$ exponents and show the results in figure \ref{fig:rivulets_9_3} of this reply.
\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{refBFig_1.pdf}
    \caption{Minkowski metric $q_2$ as a function of time for 
    $\Gamma = 1.5$ and $\lambda = 256 h_0$ from two simulations with different disjoining 
    pressure exponents, $(9,3)$ and $(3,2)$ (as in the paper) respectively.}
    \label{fig:rivulets_9_3}
\end{figure}
We have added the following footnote in the paper commenting on this aspect:\\
\\
\textcolor{red}{Expressions of 
the general form $\Pi(h) = \gamma(1-\cos \theta)(n-1)(m-1)/(n-m)((h_{\ast}/h)^n - (h_{\ast}/h)^m)/h_{\ast}$
are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending
on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
shorter. We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that 
the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801}).}
}

@@

\item[ \textbf{\underline{Comment 2.}}]
The second major concern refers to the choice of slip length. 
The authors study a fluid-substrate model that has a thin film equilibrium thickness of 
$0.07$ length units, but the slip length is one unit, hence about $10$ times larger than the equilibrium film thickness. 
This seems unexpected, in view of the small contact angle chosen for the system. 
Could the authors justify this choice?
}

\item[ \textbf{Answer}]
{
As the Referee correctly remarks, indeed small contact angles are typically associated to small slip lengths. From a phenomenological point of view, 
though, having $\delta \sim h_0$ is still commonly considered in the weak/intermediate slip regime, as proven, for instance, 
by the formation of satellite 
droplets following rivulet breakup~\cite{peschka2019signatures} or 
in the rim shape in dewetting by hole nucleation~\cite{fetzer2007quantifying, munch2005lubrication} (see also Fig.~\ref{fig:slip}).
\begin{figure}
    \centering
    \includegraphics [width=0.6\textwidth]{refBFig_2.pdf}
    \caption{Dewetting rim profiles for various slip parameters $\delta$; notice the presence of a dimple behind the front, typical of 
    small to vanishing slip, which is preserved for slip lengths up to 
    $\delta \sim 10 h_0$.}
    \label{fig:slip}
\end{figure}
Our choice of this slip length value was dictated essentially by the need 
to keep numerical stability even in the strongly time dependent case 
(large pattern speed), yet remaining in a small slip regime. 
Nevertheless, motivated by the Referee's concern and in order to check that 
picture remains qualitatively the same, we have run a simulation with 
a slip length four times smaller. The results from the analysis of the Minkowski metric $q_2$, for two cases with low and high pattern speed, reveal that indeed
the droplet-to-rivulet transition is confirmed, with $\Gamma$ as the 
discriminating parameter (see figure \ref{fig:q2smallslip}).
\begin{figure}
    \centering
    \includegraphics [width=0.75\textwidth]{refBFig_3.pdf}
    \caption{Minkowski metric $q_2$ as a function of time for two pattern speeds and slip velocity $\delta = 0.25 h_0$, signalling that the emergence of rivulets for $\Gamma>1$
    is preserved.}
    \label{fig:q2smallslip}
\end{figure}
We have commented on this when we give the value of the slip length, saying that:\\
\\
\textcolor{red}{To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
a value which lies
within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 
any major qualitative change in the results).} 
}

@@

\item[ \textbf{\underline{Comment 3.}}]
{
The relevance of the manuscript relies on the significance of the substrate's switchable properties. 
Particularly, it is assumed that the wetting properties can be modulated on space but also on time. 
The authors have mentioned in the introduction that this can be potentially done. 
However, it would be desirable that the authors relate their model of switchable substrate with some specific experimental realization, where the actual length and time scales are mentioned.
}

\item[ \textbf{{Answer}}]
{
%This is indeed a very interesting comment.
%Although our approach, as theorists/numericists, can be seen more of {\it theoretical
%engineering} and our picture would be, then, unlikely devoid of flaws, we are glad 
%to try to envisage a potential experiment.
%\JH{??}
%We thank the Referee for bringing this up.
%First, we like to note that experiments are not our field of expertise and therefore %our envisioned realization could be flawed.
%The two relevant scales are $t_0$ and $q_0$ (or $2\pi/q_0$).
%Considering the experimental parameters from~\cite{becker2003complex,PhysRevLett.99.114503} %we get, 
%\begin{align*}
%    t_0 &\approx 300\,\text{s},\qquad~~~\qquad \frac{2\pi}{q_0}\approx 
%    400\,\text{nm}.
%\end{align*}
%The time scale is in the region of minutes and the wettability pattern wavelengths are %in the interval $(800\,\text{nm}, 4000\,\text{nm})$. 
A possibility could be to prepare a thin liquid film on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
external stimuli (a light emitter). 
An ideal candidate could be a digital multimirror device (DMD).
This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
of polystyrene deposited on an oxidized silicon wafer,  
we evaluate the retraction speed 
$U_{\Theta} = \frac{\Theta^3 \gamma}{9 \mu}$ to be 
$U_{\Theta} \approx 10^{-2} \, \mu \text{m}/\text{s}$. 
The (minimum) {\it pattern speed} can be 
estimated, from the pixel size with frame rate $\sim 1 s^{-1}$, as 
$v_{\theta} \sim 10 \, \mu \text{m}/\text{s}$, that would give $\Gamma \sim 10^3$, i.e. well within 
the rivulet regime ($\Gamma > 1$). We note, also, that both $U_{\Theta}$ and $v_{\theta}$ can 
be widely modulated: the former, playing either with the (temperature and molecular weight dependent) viscosity 
or tailoring the substrate to making it more hydrophobic (i.e. increasing the contact angle); the latter, tuning
the spatial resolution and frame rate of the DMD. We expect, thus, the range of achievable $\Gamma$'s to 
be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values. 
We have included these comments in the revised manuscript just before the conclusions.

}

@@

\item[ \textbf{\underline{Comment 4.}}]
{
An important parameter in the problem is the wave-vector $q_0$, which sets the correlation length in the problem. 
The authors should state what the actual value is relative to either the average film height, or the wave-length of modulations. 
Is there any reason to give the wave-length of modulations in units of the system size, i.e., are the results subject to finite size effects? Otherwise it appears more meaningful to describe the modulations in units of $q_0$ or the equilibrium film thickness.
}

\item[ \textbf{{Answer}}]
{
We apologize for not giving the value of $q_0 \approx 0.09$ (corresponding to an intrinsic correlation length 
$\lambda_s = 2\pi/q_0 \approx 70 h_0$), which is now provided in the revised manuscript. We also acknowledge that expressing the wavelength in terms of the size $L$
was not at all meaningful (no finite size effects are involved). In fact, all the lengths (including $L$) are expressed in units of the mean film thickness $h_0$ (and it was so also in the old version of the paper, as commented in the text below eq. (4)). In order to avoid any confusion, we explicitly write 
the values of all lengths in terms of $h_0$ in the revised version.
}

\item[ \textbf{\underline{Comment 5.}}]
{
The thin film equation relies on a local approximation for the interface potential and disjoining pressure. 
Notice that to first order in non-local effects, the surface tension entering the thin film equation adopts a film thickness dependence which could potentially affect the dynamics. 
Particularly, as far as the parallel correlation length $q_0$ is referred, the non-local effects provide an upper bound equal to the inverse bulk correlation length. 
This could be a significant effect in polymer thin films (c.f. Benet et al. jpc-c 118, 22079, 2014 and the related discussion on dynamical effects Pahlavan et al., jfm, 845, 642, 2018).
}

\item[ \textbf{{Answer}}]
{
We guess that the Referee mentions the fact that the Laplacian of the thickness field, $\nabla^2 h$,
is just a small gradient approximation of the actual local curvature, such that the Laplace contribution to the film pressure should indeed read $\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$. This is of course correct, but 
the condition $|\nabla h| \ll 1$ holds true in our system, owing to the very same lubrication approximation,
or, equivalently, to the small contact angles involved; the spinodal wavelength for a homogeneous substrate with
contact angle $\theta_0$ is $\lambda_s = 2\pi/q_0(\theta_0) \approx 70 h_0$. Therefore, the typical gradients
of the thickness field can be estimated as $|\nabla h| \sim h_0/\lambda_s \approx 1/70 \ll 1$ (notice 
that this is, somehow, a worst case scenario since, as we discuss in the text, in the patterned substrates
$\lambda \stackrel{>}{\sim} \lambda_s$).
We have added the following footnote, where we introduce the film pressure:\\
\\
\textcolor{red}{The Laplace pressure is the product of the surface tension 
and the local curvature, therefore its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$; 
since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
and the pressure reduces to the form involving only the Laplacian of the thickness field~\cite{Benet2014,alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}}, \\
\\
citing the related papers as a reference.
}
\end{itemize}

### 2022-12-06 Old_version/main_AS.tex
First resubmission

\author{S. Zitz}
\email{zitz@ruc.dk}
 %Changed order: Work done while Stefan worked in Nuremberg
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
  \affiliation{IMFUFA, Department of Science and Environment,\\ 
Roskilde University, Postbox 260, DK-4000 Roskilde, Denmark}%
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), 
Consiglio Nazionale delle Ricerche (CNR),\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

@@

\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness.

@@

@@

$L/h_0 = 512 $.
To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$ (see Eqs.~(\ref{eq:mobility})-(\ref{eq:disjoinpressure})). 

@@

    \includegraphics[width=0.45\textwidth]{Figures/low_lambda_rupture.pdf}

@@

    \includegraphics[width=0.45\textwidth]{Figures/static_pattern_test.pdf}

@@

$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, $\lambda = \{L/3,L/2,L \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines), showing excellent agreement.\\

@@

    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values (see Eq.~(\ref{eq:vel_ratio})), on a substrate with pattern wavelength $\lambda=L/2$.

>>>>

\author{S. Zitz}https://www.overleaf.com/project/633b184a2c25f373c2e9f56c
\email{zitz@ruc.dk}
 %Changed order: Work done while Stefan worked in Nuremberg
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
  \affiliation{IMFUFA, Department of Science and Environment,\\ 
Roskilde University, Postbox 260, DK-4000 Roskilde, Denmark}%
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), 
Consiglio Nazionale delle Ricerche (CNR),\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

@@

\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness~\footnote{\textcolor{red}{Expressions of 
the general form $\Pi(h) = \gamma(1-\cos \theta)(n-1)(m-1)/(n-m)((h_{\ast}/h)^n - (h_{\ast}/h)^m)/h_{\ast}$
are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending
on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
shorter. We have run also a single case ($\lambda=256$ and $\Gamma=15$) with $(9,3)$ and observed that 
the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801})}}.

@@

%%%%
\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
 $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
 the Hamaker constant $A_H$. Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 
dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
The characteristic time of the growth of the instability for these experiments, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$) $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ and it can be seen that indeed film rupture occurs within time 
scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}, as in our 
simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
with homogeneous substrates~\cite{PhysRevE.104.034801}).}~\footnote{\textcolor{red}{A laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
The external stimulus, our moving pattern, can be realized by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}}
%%%%

@@

$L/h_0 = 512 $. \textcolor{red}{The numerical values in lattice Boltzmann units of the 
mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
and $\gamma=0.01$.}
\textcolor{red}{To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
a value which lies
within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 
any major qualitative change in the results).} 

@@

    \includegraphics[width=0.45\textwidth]{Gamma_rupT_with_scaling.pdf}

@@

    \includegraphics[width=0.45\textwidth]{Figures/static_pattern_test_no_hd.pdf}

@@

$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, 
$\lambda/h_0 = \{170,256,512 \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines), showing excellent agreement.\\

@@

    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values (see Eq.~(\ref{eq:vel_ratio})), on a substrate with pattern wavelength $\lambda=256 h_0$.
